\EL{Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques}

### 2020-02-27 jfm/draft_p2_jfm.tex
Relecture apres Emmanuel

\begin{abstract}
	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a  turbulent flow impinging onto a bluff body, and examines the relevance of two distinct statistical algorithms to efficiently sample these events. 
	%
	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
	%
	Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle. This vortex is responsible for a significant pressure drop between the forebody and the base of the obstacle, thus yielding a very high value of the drag. 
	%
	Two algorithms are then considered to speed up the sampling of such flow scenarii related to extreme fluctuations of the drag, namely the \ac{ams} and the \ac{gktl} algorithms. 
	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned (according to some specific rules) to sample large-amplitude events more frequently.  
	%

		\QQ{Justifier le choix de ces deux algorithmes et mentionner les différences et complémentarités ?}
		%\ZZ{These latter are representative of the two main classes of algorithms applicable to deterministic out-of-equilibrium dynamics. Whilst the \ac{ams} is general, the \ac{gktl} algorithm is suited to sample trajectories based on the time-averaged value of the observable.}

	%
	These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry, but it is here the first application to a fluid-structure interaction problem. 
	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
	%
	Practical evidence is given that the fast sweeping time of fluid structures past the obstacle has a strong influence on the efficiency of the \QQ{selection-mutation or selection-cloning (the terminology should be unified in the whole paper)} rules of these two algorithms. 
	% 
	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
	%
	\QQ{phrase de conclusion ?}
}
%\sout{This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
%The main idea is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently. 
%%
%Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events otherwise out of reach by direct sampling.
%This work is the first application to fluid-structure interaction problems. 
%%
%The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major rare-event sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
%Practical evidence is given that the fast sweeping-time of fluid structures past the obstacle has a drastic influence on the efficiency of these two algorithms.
%While it is shown that the \ac{ams} algorithm does not yield significant run-time savings, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times. 
%%
%Beyond the study of applicability of rare-event sampling techniques to a fluid-mechanical problem, this work also includes a detailed phenomenological description of extreme-drag events of a turbulent flow on a bluff body. }
\end{abstract}

@@

From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \citep{KRAICHNAN}.
The use of analytical perturbative methods in identifying resonant interactions (among degrees of freedom) responsible for extreme fluctuations is unsuccessful.
Alternatively, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
However, this requires very long simulations since these extreme events are rare. \EL{The computational cost for sampling a fluctuation of very small probability typically grows as the inverse of this probability~\citep{wouters2016rare}.}
%
%However, this requires very long simulations since these events are rare.
% general def on rare-event sampling %
%
%
{Rare-event sampling \EL{refers} to a large body of methods that aim at \emph{preferentially} exploring  regions of phase space corresponding to  events that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
%
%	
% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
%
In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is \EL{conducted} by using both very long time-series (direct sampling) and rare-event sampling techniques.

%
In fluid turbulence, rare-event sampling has been approached mainly from the perspective of simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \citep{bec_burgers_2007}. In this case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \citep{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
%
An alternative approach is based on instantons \citep{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
Instantons refer to the most probable trajectories in phase space that achieve a given rare event (in the limit of weak noise). Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \citep{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
An example is the investigation of the physics of rogue waves~\citep{dematteis2018rogue,dematteis2019experimental}.
%
However, a drawback of the aforementioned approaches is their limitation to simple and stochastically driven dynamics.
%
%
%
Here, a more general approach is considered for complex, possibly deterministic, dynamical systems.
It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories of the system. 
%, and is designed to sample rare events of some observable with a higher frequency.
%    
%
Even though such ideas date back to the early 1950s, they have received ever-growing interest over the the last twenty years with successful applications in various domains such as chemistry \citep{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \citep{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \citep{louvin2017}, nonlinear dynamical systems \citep{tailleur_probing_2007} and communication networks simulation \citep{villen-altamirano_restart:_1994}.
More importantly, these types of algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\citep{wouters2016rare}.
%
%
\EL{An original contribution of our study is certainly to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom. 
%
Two algorithms a priori suitable for such dynamics are considered, namely, the Adaptive Multilevel Splitting (AMS) algorithm \citep{cerou_adaptive_2007} and the Giardina-Kurchan-Tailleur-Lecomte (GKTL) algorithm \citep{giardina_direct_2006}.}
%

@@

% Annonce du plan
In more details,  the flow set-up is introduced and the dynamics related to typical
drag fluctuations is described in section~\ref{sec:test_flow}.
The statistical properties of the drag are then discussed.
In section~\ref{sec:direct_sampling}, \EL{the phenomenology of  extreme drag fluctuations} is investigated based on direct sampling.
Both the instantaneous drag and time-averaged drag are considered.
It is found in particular that sampled extreme events for the instantaneous drag result from very similar dynamics. %Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
%\sout{This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.}
Section~\ref{sec:rare_events_algorithms} examines the applicability of both the \ac{ams} and \ac{gktl} algorithms to the simulation of extreme drag fluctuations in the same flow configuration.
In subsection~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforwardly. This difficulty is put in perspective with the phenomenology of extreme drag fluctuations developed in  section \ref{sec:direct_sampling}.
Subsection~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag by using the \ac{gktl} algorithm.
This latter allows for an exceptional reduction of the computational cost needed to simulate trajectories corresponding to extreme values of the time-averaged drag.
As a specific \EL{successful} application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
Perspectives and conclusion end this work.

@@

% lattice Boltzmann method 
%
The flow dynamics is integrated numerically by the \ac{lbm} \EL{\citep{book_lbm}}. 
While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at an underlying  mesoscopic level.
Capturing the dynamics of collections of fluid particles moving and colliding on a lattice is here preferred to solving non-linear PDEs. 
%This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level. Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
Further details about the \ac{lbm} are given in Appendix \ref{app:lbm} and references therein.
In our context, this numerical method has been chosen principally for its computational efficiency.

% geometry
%
The simulated flow develops in a long plane channel of dimension $513 \times 129$ mesh points. The square obstacle has size $R=16$ (in mesh unit) and is located at the centre of the channel. The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}). 
% boundary conditions
%
No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure~\citep{lbm_book}.  
%
Upstream of the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as an inlet condition. 
The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively. The initial distributions are imposed at equilibrium (see Appendix \ref{app:lbm}). 
In the bulk, the viscosity is adjusted so that grid turbulence is generated with Reynolds number $\mathrm{Re_{grid}}=1200$. The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the \ac{lbm}. 
Near the end of the channel, the flow is progressively damped within a \textit{sponge zone} where the viscosity is artificially enhanced.
Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \cite{latt2008straight}.

@@

% drag statistics
% pdf
Drag fluctuations have been sampled along a simulation of duration \EL{$T_{tot} = 4\times 10^6~T_0$}.
This long simulation will be referred to as the \textit{control run} in the following.
It has been made possible by the relative simplicity of the investigated flow and the computational efficiency of the lattice Boltzmann method. 
The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  \EL{${\mathbb{P}}(f_d) \propto e^{-\ell f_d}$}.
%
Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface corresponding to the periphery of the obstacle but in the absence of the obstacle. 
%
In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails. This shows that the asymmetry of the \ac{pdf} and the development of a positive exponential tail are closely related to the no-slip condition on the obstacle boundary.
\begin{figure}
	\centering
	\subfloat[\ac{pdf} of normalised drag fluctuations]
	{\label{fig:pdf_drag_a}
		\includegraphics[width=.45\linewidth]{./PDF_drag/PDF_drag.png}}
	\subfloat[Autocorrelation of drag fluctuations]
	{\label{fig:pdf_drag_b}
		\includegraphics[width=.45\linewidth]{autocorrelation_drag/autocorrelation_drag.png}}  
	\caption{{(a)} \ac{pdf} of normalised drag fluctuations ($\tilde f_d = f_d - \bar{f_d}$) where $\mu = \bar{f_d}$ denotes the time-averaged value and $\sigma = \sqrt{\overline{{\tilde f_d}^2}}$ is the standard deviation. The drag is evaluated both in the presence (blue) and in the absence (red) of the obstacle. %Note that the amplitudes have not been normalized.
		\QQ{x-axis should be $x = (f_d-\mu)/\sigma)$ and y-axis $P(x)$)}
		{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$. The correlation time \EL{$\tau_c\simeq 4 T_0$} is given by $C(\tau_c)=0$.
		%$\tau_c$ may therefore be considered as the correlation time of the drag signal.
	}
	\label{fig:pdf_drag}
\end{figure}

@@

How rare is a fluctuation $a$ is quantified by the return time $r(a)$.
We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
%
%
%The shape of the return time $r$ as a function of the return level $a$ is system dependent. 
If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
% As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag})
one might expect that 
\begin{equation}
\label{eq:return_time}
\EL{r(a) \underset{a\to\infty}{\propto} \exp(la)}
\end{equation}
where $l$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$~\citep{lestang_computing_2018}.
Consistently, it is found that the return time $r(a)$ is well approximated by an exponential for large levels $a$. Let us also point out some deviation from the exponential law at the largest levels, which is probably the consequence of under-sampling.

@@

% intro
%
We have extracted the fluctuations of the drag with a return time $r(a)$ greater than  $10^4\tau_c$ from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
This set will be considered as representative of \emph{extreme events} in the upcoming study.
The choice of this particular threshold has been driven by the need to collect enough events with large amplitude and to possibly identify generic features.
%
According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma =\sqrt{\overline{\tilde{f_d}^2}}$ being the standard deviation of the drag process.
Precisely, 104 independent fluctuations with $f_d(t) \geq 7.6\sigma$ have been identified. Each fluctuation is characterized by its maximal value, $f_d^{\star}$, and the time, $t^{\star}$, at which this maximum is reached.
%
In the following, the phenomenology of extreme drag fluctuations will be examined on the basis of this set of events.

@@

% teasing
%
%\sout{In section~\ref{sec:test_flow}, it was pointed out that typical drag fluctuations originate mostly from the variation of the forebody pressure, \textit{i.e.} from the upstream turbulent flow. We shall see that the situation is different in the case of {extreme} drag fluctuations.} 
\EL{In section~\ref{sec:test_flow}, it was pointed out that drag fluctuations within one standard deviation were not associated with any particular arrangement of the   vorticity around the obstacle. We shall see in the following that the situation is different in the case of {extreme} drag fluctuations. }

% relative contribution of forebody and base pressure
%
Let $(t^{\star}, f_d^{\star})$ refer to an extreme-drag event.
The (zero-mean) fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \overline{f_d}$ can be  decomposed into 
\begin{equation}
\tilde{f}_d^{\star} = \Delta p_{fb}^{\star} - \Delta p_{base}^{\star}
\end{equation}
where $\Delta p_{fb}^{\star}$ and $\Delta p_{base}^{\star}$ denote the variations of the forebody and base pressure, respectively.
%
Fig.~\ref{fig:pressure_ratio} displays the relative contributions
$\Delta p_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\Delta p_{base}^{\star}/\tilde{f}_d^{\star}$ to \EL{the drag fluctuation $\tilde f_d^{\star}$.
	It is found that very large fluctuations of the drag result mainly ($\sim 80\%$) from the drop of the base pressure, whereas the variation of forebody pressure contributes much less to $\tilde{f}_d^{\star}$. On the contrary, moderate fluctuations arise from combined variations of the forebody and base pressures without any particular predominance, which is in agreement with the previous observations (see Fig.\ref{fig:typical_vorticity}).}

@@

s

@@

\newpage

>>>>

\begin{abstract}
	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a turbulent flow impinging on a bluff body, and examines the relevance of two distinct rare-event algorithms to efficiently sample these events. 
	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
	Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle. This vortex is responsible for a significant pressure drop between the forebody and the base of the obstacle, thus yielding a very high value of the drag. 
	Two algorithms are then considered to speed up the sampling of such flow scenarii, namely the \ac{ams} and the \ac{gktl} algorithms. 
	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned, according to some specific rules designed to sample large-amplitude events more frequently.  These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry.
        The present study is the first application to a fluid-structure interaction problem. 
	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
	Practical evidence is given that the fast sweeping time of turbulent fluid structures past the obstacle has a strong influence on the efficiency of the rare-event algorithm. 
	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be effective to sample very efficiently extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.}
%	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a  turbulent flow impinging onto a bluff body, and examines the relevance of two distinct statistical algorithms to efficiently sample these events. 
%	%
%	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
%	%
%	Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle. This vortex is responsible for a significant pressure drop between the forebody and the base of the obstacle, thus yielding a very high value of the drag. 
%	%
%	Two algorithms are then considered to speed up the sampling of such flow scenarii related to extreme fluctuations of the drag, namely the \ac{ams} and the \ac{gktl} algorithms. 
%	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned (according to some specific rules) to sample large-amplitude events more frequently.  
%	%
%		\QQ{Justifier le choix de ces deux algorithmes et mentionner les différences et complémentarités ?}
%		%\ZZ{These latter are representative of the two main classes of algorithms applicable to deterministic out-of-equilibrium dynamics. Whilst the \ac{ams} is general, the \ac{gktl} algorithm is suited to sample trajectories based on the time-averaged value of the observable.} 
%	%
%	These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry, but it is here the first application to a fluid-structure interaction problem. 
%	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
%	%
%	Practical evidence is given that the fast sweeping time of fluid structures past the obstacle has a strong influence on the efficiency of the \QQ{selection-mutation or selection-cloning (the terminology should be unified in the whole paper)} rules of these two algorithms. 
%	% 
%	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
%	%
%	\QQ{phrase de conclusion ?}
%\sout{This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
%The main idea is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently. 
%Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events otherwise out of reach by direct sampling.
%This work is the first application to fluid-structure interaction problems. 
%The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major rare-event sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
%Practical evidence is given that the fast sweeping-time of fluid structures past the obstacle has a drastic influence on the efficiency of these two algorithms.
%While it is shown that the \ac{ams} algorithm does not yield significant run-time savings, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times. 
%Beyond the study of applicability of rare-event sampling techniques to a fluid-mechanical problem, this work also includes a detailed phenomenological description of extreme-drag events of a turbulent flow on a bluff body. }
\end{abstract}

@@

From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \citep{KRAICHNAN}.
The use of analytical perturbative methods in identifying resonant interactions (among degrees of freedom) responsible for extreme fluctuations is unsuccessful.
Alternatively, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
However, this requires very long simulations since these extreme events are rare. \EL{The computational cost for sampling a fluctuation of very small probability typically grows as the inverse of this probability~\citep{wouters2016rare}.}
%
%However, this requires very long simulations since these events are rare.
% general def on rare-event sampling %
%
%
{Rare-event sampling \EL{refers} to a large body of methods that aim at \emph{preferentially} exploring  regions of phase space corresponding to  events that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
%
%	
% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
%
In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is \EL{conducted} using both very long time-series (direct sampling) and rare-event sampling techniques.

%
In fluid turbulence, rare-event sampling has been approached mainly from the perspective of simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \citep{bec_burgers_2007}. In this case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \citep{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
%
An alternative approach is based on instantons \citep{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
Instantons refer to the most probable trajectories in phase space that achieve a given rare event (in the limit of weak noise). Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \citep{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
An example is the investigation of the physics of rogue waves~\citep{dematteis2018rogue,dematteis2019experimental}.
%
However, a drawback of the aforementioned approaches is their limitation to simple and stochastically driven dynamics.
%
%
%
Here, a more general approach is considered for complex, possibly deterministic, dynamical systems.
It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories of the system. 
%, and is designed to sample rare events of some observable with a higher frequency.
%    
%
Even though such ideas date back to the early 1950s, they have received growing interest over the last twenty years with successful applications in various domains such as chemistry \citep{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \citep{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \citep{louvin2017}, nonlinear dynamical systems \citep{tailleur_probing_2007} and communication networks simulation \citep{villen-altamirano_restart:_1994}.
More importantly, such algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\citep{wouters2016rare}.
%
%
\EL{Certainly, an original contribution of our study is to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom. 
%
Two algorithms that are a priori suitable for such dynamics are considered, namely, the Adaptive Multilevel Splitting (AMS) algorithm \citep{cerou_adaptive_2007} and the Giardina-Kurchan-Tailleur-Lecomte (GKTL) algorithm \citep{giardina_direct_2006}.}
%

@@

% Annonce du plan
The flow set-up is introduced and the dynamics related to typical
drag fluctuations is described in section~\ref{sec:test_flow}.
The statistical properties of the drag are then discussed.
In section~\ref{sec:direct_sampling}, \EL{the phenomenology of  extreme drag fluctuations} is investigated based on direct sampling.
Both the instantaneous drag and time-averaged drag are considered.
It is found in particular that sampled extreme fluctuations of the instantaneous drag result from very similar dynamics. %Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
%\sout{This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.}
Section~\ref{sec:rare_events_algorithms} examines the applicability of both the \ac{ams} and \ac{gktl} algorithms to the simulation of extreme drag fluctuations in the same flow configuration.
In subsection~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforwardly. This difficulty is put in perspective with the phenomenology of extreme drag fluctuations developed in  section \ref{sec:direct_sampling}.
Subsection~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag by using the \ac{gktl} algorithm.
This latter allows for an exceptional reduction of the computational cost needed to simulate trajectories corresponding to extreme values of the time-averaged drag.
As a specific \EL{successful} application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
Perspectives and conclusion end this work.

@@

% lattice Boltzmann method 
%
The flow dynamics is integrated numerically by the \ac{lbm} \EL{\citep{kruger_lattice_2017}}. 
While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at an underlying  mesoscopic level.
Capturing the dynamics of collections of fluid particles moving and colliding on a lattice is here preferred to solving non-linear PDEs. 
%This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level. Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
Further details about the \ac{lbm} are given in Appendix \ref{app:lbm} and references therein.
In our context, this numerical method has been chosen principally for its computational efficiency.

% geometry
%
The simulated flow develops in a long plane channel of dimension $513 \times 129$ mesh points. The square obstacle has size $R=16$ (in mesh units) and is located at the centre of the channel. The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}). 
% boundary conditions
%
No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure~\citep{lbm_book}.  
%
Upstream of the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as an inlet condition. 
The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively. The initial distributions are imposed at equilibrium (see Appendix \ref{app:lbm}). 
In the bulk, the viscosity is adjusted so that grid turbulence is generated with Reynolds number $\mathrm{Re_{grid}}=1200$. The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the \ac{lbm}. 
Near the end of the channel, the flow is progressively damped within a \textit{sponge zone} where the viscosity is artificially enhanced.
Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \citep{latt2008straight}. For more details about the implementation of the flow, see the software repository attached to this paper \textbf{?????????}.

@@

% drag statistics
% pdf
Drag fluctuations have been sampled along a simulation of duration \EL{$T_{tot} = 4\times 10^6~T_0$}.
This long simulation will be referred to as the \textit{control run} in the following.
It has been made possible by the relative simplicity of the investigated flow and the computational efficiency of the \ac{lbm}.
The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  \EL{${\mathbb{P}}(f_d) \propto e^{-\ell f_d}$}.
%
Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface corresponding to the periphery of the obstacle but in the absence of the obstacle. 
%
In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails. This shows that the asymmetry of the \ac{pdf} and the development of a positive exponential tail are closely related to the no-slip condition on the obstacle boundary.
\begin{figure}
	\centering
	\subfloat[\ac{pdf} of normalised drag fluctuations]
	{\label{fig:pdf_drag_a}
		\includegraphics[width=.4\linewidth]{./PDF_drag/PDF_drag.png}}
	\subfloat[Autocorrelation of drag fluctuations]
	{\label{fig:pdf_drag_b}
		\includegraphics[width=.45\linewidth]{autocorrelation_drag/autocorrelation_drag.png}}  
	\caption{{(a)} \ac{pdf} of normalised drag fluctuations ($\tilde f_d = f_d - \bar{f_d}$) where $\mu = \bar{f_d}$ denotes the time-averaged value and $\sigma = \sqrt{\overline{{\tilde f_d}^2}}$ is the standard deviation. The drag is evaluated both in the presence (blue) and in the absence (red) of the obstacle. %Note that the amplitudes have not been normalized.
		{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$. The correlation time \EL{$\tau_c\simeq 4 T_0$} is given by $C(\tau_c)=0$.
		%$\tau_c$ may therefore be considered as the correlation time of the drag signal.
	}
	\label{fig:pdf_drag}
\end{figure}

@@

How rare is a fluctuation $a$ is quantified by the return time $r(a)$.
We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
%
%
%The shape of the return time $r$ as a function of the return level $a$ is system dependent. 
If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
% As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag})
one might expect that 
\begin{equation}
  \label{eq:return_time}
  \EL{r(a) \underset{a\to\infty}{\propto} \exp(\ell a)}
\end{equation}
where $l$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$~\citep{lestang_computing_2018}.
Consistently, it is found that the return time $r(a)$ is well approximated by an exponential for large levels $a$. Let us also point out some deviation from the exponential law at the largest levels, which is probably the consequence of under-sampling.

@@

% intro
%
We have extracted the fluctuations of the drag with a return time $r(a)$ greater than  $10^4\tau_c$ from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
This set will be considered as representative of \emph{extreme events} in the upcoming study.
The choice of this particular threshold has been driven by the need to collect enough events with large amplitude and to possibly identify generic features.
%
According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma =\sqrt{\overline{\tilde{f_d}^2}}$ being the standard deviation of the drag process, where $\tilde{f_d}$ denotes the drag with zero-mean, \textit{i.e.} $\tilde{f}_d = f_d - \overline{f_d}$.
Precisely, 104 independent fluctuations with $\tilde{f}_d(t) \geq 7.6\sigma$ have been identified. Each fluctuation is characterized by its maximal value, $f_d^{\star}$, and the time, $t^{\star}$, at which this maximum is reached.
%
In the following, the phenomenology of extreme drag fluctuations will be examined on the basis of this set of events.

@@

% teasing
%
%\sout{In section~\ref{sec:test_flow}, it was pointed out that typical drag fluctuations originate mostly from the variation of the forebody pressure, \textit{i.e.} from the upstream turbulent flow. We shall see that the situation is different in the case of {extreme} drag fluctuations.} 
\EL{In section~\ref{sec:test_flow}, it was pointed out that drag fluctuations within one standard deviation were not associated with any particular arrangement of the   vorticity around the obstacle. We shall see in the following that the situation is different in the case of {extreme} drag fluctuations. }
%
% relative contribution of forebody and base pressure
%
Let $(t^{\star}, f_d^{\star})$ refer to an extreme-drag event.
The (zero-mean) fluctuation $\tilde{f}_d^{\star} = f_d^{\star} - \overline{f_d}$ can be  decomposed into 
\begin{equation}
\tilde{f}_d^{\star} = \Delta p_{fb}^{\star} - \Delta p_{base}^{\star}
\end{equation}
where $\Delta p_{fb}^{\star}$ and $\Delta p_{base}^{\star}$ denote the variations of the forebody and base pressure, respectively.
%
Fig.~\ref{fig:pressure_ratio} displays the relative contributions
$\Delta p_{fb}^{\star}/\tilde{f}_d^{\star}$ and $-\Delta p_{base}^{\star}/\tilde{f}_d^{\star}$ to \EL{the drag fluctuation $\tilde f_d^{\star}$.
	It is found that very large fluctuations of the drag result mainly ($\sim 80\%$) from a drop in base pressure, whereas the variation of forebody pressure contributes much less to $\tilde{f}_d^{\star}$. On the contrary, moderate fluctuations arise from combined variations of the forebody and base pressures without any particular predominance, which is in agreement with the previous observations (see Fig.\ref{fig:typical_vorticity}).}

@@

@@
